Whittaker Models for Real Groups

### 2024-04-25 whittaker.tex
more changes in progress

Suppose $\pi$ is an irreducible representation of~$G(\R)$.
We say $\pi$ has a {\it Whittaker model}, or is {\it generic},  if
$\pi$ embeds into the induced
representation $\mathrm{Ind}_{N(\R)}^{G(\R)}(\chi)$ for some $(B,\eta)$ satisfying:
We consider pairs
\begin{enumerate}[(i)]
\item $B$ is a Borel subgroup of~$G$ defined over~$\R$, with real points $B(\R)$,
\item $\eta$ is a character of the nilpotent radical~$N(\R)$ of~$B(\R)$, subject to the condition that~$\eta$ is nontrivial on each simple root space.
\end{enumerate}
The existence of a Borel subgroup satisfying (i) is the definition of $G$ being quasisplit, which we assume holds from now on.
Clearly this condition only depends on the conjugacy class of the pair
so we refer to consider pairs $(B,\eta)$ modulo conjugation by $G(\R)$, and refer to such a conjugacy class as a 
as a {\it Whittaker datum}.
This is a small finite set.
If $\w$ is a Whittaker datum, it makes sense to talk about $\pi$ having a $\w$-Whittaker model. We let $\Wh(\pi)$  be the set of
Whittaker data $\w$ such that $\pi$ has a $\w$-Whittaker model.

@@

We work main in the setting of $(\g,K)$-modules.

@@

We work mostly in the  algebraic setting of
$(\g,K)$-modules. However, mainly for the analytic wave front set
invariant we need to consider representations of $G(\R)$ on a Banach
space. Suppose $(\pi,V)$ is an irreducible representation of $G(\R)$ 
 where $V$ is a Banach space. The underlying
$(\g,K)$-module is on the space of $K$-finite vectors in
$V$. Conversely if $\pi$ is an irreducible $(\g,K)$-module it can be
realized as the $K$-finite vectors of some  Banach space
representation.

@@

Now suppose $(\pi,V)$ is an irreducible representation of $G(\R)$. 
See \cite{howe_wave_front}, \cite{bv_local_structure},
\cite{HarrisHeOlafsson} for the definition of the \emph{wave-front
  set} $\WF(\pi)$. This is an analytic invariant determined by the
singularities of the global character of $\pi$ near the identity. 
It is a closed $G(\R)$-invariant subset of the nilpotent cone in
$i\g(\R)^*$. As such it is the union of a finite number of nilpotent $G(\R)$-orbits.

@@

Based on the Lemma we define $Q(G)$ to be $Q_\sigma(G)$ or $Q_\theta(G)$, depending on the setting.

@@

>>>>

We consider pairs
\begin{enumerate}[(i)]
\item $B$ is a Borel subgroup of~$G$ defined over~$\R$, with real points $B(\R)$,
\item $\eta$ is a character of the nilpotent radical~$N(\R)$ of~$B(\R)$, subject to the condition that~$\eta$ is nontrivial on each simple root space.
\end{enumerate}
The existence of a Borel subgroup satisfying (i) is the definition of $G$ being quasisplit, which we assume holds from now on.

We recall what it means for a representation $\pi$ to have a $(B,\eta)$-Whittaker model in Section ?.
This condition only depends on the conjugacy class of the pair, 
so we  consider pairs $(B,\eta)$ modulo conjugation by $G(\R)$, and refer to such a conjugacy class as a 
as a {\it Whittaker datum}.
This is a small finite set.
If $\w$ is a Whittaker datum, it makes sense to talk about $\pi$ having a $\w$-Whittaker model. We let $\Wh(\pi)$  be the set of
Whittaker data $\w$ such that $\pi$ has a $\w$-Whittaker model. We say $\pi$ is {\it generic} if $\Wh(\pi)$ is non-empty.

@@

@@

We need to pass back and forth between representations of $G(\R)$ and $(\g,K)$-modules. 
See ? for details.

Suppose $(\pi,V)$ is an irreducible representation of $G(\R)$  on a Hilbert space. 
The underlying
$(\g,K)$-module is on the space of $K$-finite vectors in
$V$. Conversely if $\pi$ is an irreducible $(\g,K)$-module it can be
realized as the $K$-finite vectors of a Hilbert space representation, which we 
refer to as the Hilbert space globalization of $\pi$.

@@

Now suppose $(\pi,V)$ is an irreducible representation of $G(\R)$ on a Hilbert space. 
See \cite{howe_wave_front}, \cite{bv_local_structure},
\cite{HarrisHeOlafsson} for the definition of the \emph{wave-front
  set} $\WF(\pi)$. This is an analytic invariant determined by the
singularities of the global character of $\pi$ near the identity. 
It is a closed $G(\R)$-invariant subset of the nilpotent cone in
$i\g(\R)^*$. As such it is the union of a finite number of nilpotent $G(\R)$-orbits.

If $\pi$ is an irreducible $(\g,K)$-module we define $\WF(\pi)$ to be the wave-front set of 
the Hilbert space globalization of $\pi$. 
In the other direction if $\pi$ is an irreducible representation of $G(\R)$ we define $\AV(\pi)$ to be the associated variety 
of its underlying $(\g,K)$-module.

[Need a paragraph explaining $\Wh(\pi)$, following Matsumoto]

@@

@@

Note that $Q_\sigma(G)$ is a set of automorphisms of $G(\R)$. It therefore acts on representations of $G(\R)$. 
Similarly $Q_\theta(G)$ is a set of automorphisms of $(\g,K)$, and acts on $(\g,K)$-modules.

\begin{lemma}
Suppose $\pi$ is an irreducible  Hilbert space representation of $G(\R)$.
Suppose $g\in Q_\sigma(G)$. 
Then $(\pi^g)_K\simeq (\pi_K)^{\phi(g)}$.
\end{lemma}

\begin{proof}
straightforward
\end{proof}

Based on these two Lemmas we write $Q(G)$ for the group $Q_\sigma(G)$ or $Q_\theta(G)$, depending on the situation.
